\documentclass[11pt,a4paper]{report}

\usepackage[a4paper,margin=24mm,headheight=14pt]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{array}
\usepackage{tabularx}
\usepackage{float}
\usepackage{caption}
\usepackage{fancyhdr}
\usepackage[hidelinks]{hyperref}
\usepackage{enumitem}
\usepackage{url}

\definecolor{ink}{RGB}{32,44,58}
\definecolor{accent}{RGB}{37,96,151}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\setlist[itemize]{leftmargin=18pt,itemsep=2pt,topsep=3pt}
\setlist[enumerate]{leftmargin=18pt,itemsep=2pt,topsep=3pt}
\renewcommand{\arraystretch}{1.14}
\captionsetup{font=small,labelfont=bf}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\textcolor{ink}{Model Inclusion Criteria}}
\fancyhead[R]{\textcolor{ink}{\thepage}}
\renewcommand{\headrulewidth}{0.4pt}

\newcommand{\code}[1]{\texttt{#1}}

\title{\textbf{Seeing the Same Task Differently}[8pt]
Vision-Language Model Inclusion Criteria}
\author{Technical report from the repository evaluation outputs}
\date{}

\begin{document}
\maketitle

\chapter{Model Inclusion Criteria}

\section{Purpose of the Selection}

This study evaluates a deliberately bounded set of vision-language models. It
does not attempt to reproduce a full public leaderboard or catalogue every
model available at the time of execution. The purpose is narrower: to compare a
small group of practically accessible, general-purpose multimodal systems under
one benchmark implementation.

The model choice was informed by two contemporary vision-language leaderboards:
MMMU-Pro\footnote{\url{https://artificialanalysis.ai/evaluations/mmmu-pro}, accessed 2026-06-11.}
and Arena Vision Overall\footnote{\url{https://arena.ai/leaderboard/vision}, accessed 2026-06-11.}.
The MMMU-Pro table below uses the public LLM Stats MMMU-Pro leaderboard
snapshot.\footnote{\url{https://llm-stats.com/benchmarks/mmmu-pro}, accessed 2026-06-11.}
These leaderboards identify relevant model families; the repository artifacts
determine which models can actually be compared in this study.

The final cohort contains six models:

\begin{itemize}
\item \textbf{OpenAI}: GPT-4o, GPT-4.1, GPT-5;
\item \textbf{Anthropic}: Claude Sonnet 4.5;
\item \textbf{Google}: Gemini 3.5 Flash;
\item \textbf{Alibaba Cloud}: Qwen 3.5.
\end{itemize}

This selection gives extra depth to the GPT family because it is the strongest
and most consistently relevant family for this study: GPT models appear
repeatedly near the top of the leaderboard evidence, and the repository
contains multiple completed GPT-family artifacts under the same benchmark
pipeline. The other major families are represented by one model each, which
keeps the comparison broad without turning the study into a full leaderboard
survey.

\section{Why These Two Leaderboards Were Used}

MMMU-Pro and Arena Vision Overall were chosen because they provide complementary
views of current vision-language performance. MMMU-Pro emphasizes structured
multimodal reasoning, while Arena Vision Overall reflects broader comparative
model behavior across vision-language tasks. Using both reduces dependence on a
single benchmark source while keeping the selection process compact.

The leaderboard excerpts in Tables~\ref{tab:mmmu-pro-context}
and~\ref{tab:arena-vision-context} are fixed snapshots recorded at the time of
model selection. Each table contains 25 entries: enough to show the main
families visible near the top of each ranking, but not so many that the
inclusion argument becomes a leaderboard survey.

\section{MMMU-Pro Leaderboard Context}

Table~\ref{tab:mmmu-pro-context} presents the MMMU-Pro leaderboard context used
during model selection. It shows that GPT, Gemini, Qwen, Claude, and Muse Spark
all belong to the relevant upper part of the model landscape.

\begin{table}[H]
\centering
\small
\resizebox{\textwidth}{!}{%
\begin{tabular}{l c c l l}
\toprule
\textbf{Model name} &
\textbf{Rank} &
\textbf{Score} &
\textbf{Family} &
\textbf{Company} \\
\midrule
Gemini 3.5 Flash
& 1 & 0.836 & Gemini & Google \\
GPT-5.5
& 2 & 0.832 & GPT & OpenAI \\
Gemini 3 Flash
& 3 & 0.812 & Gemini & Google \\
GPT-5.4
& 3 & 0.812 & GPT & OpenAI \\
Gemini 3 Pro
& 5 & 0.810 & Gemini & Google \\
Gemini 3.1 Pro
& 6 & 0.805 & Gemini & Google \\
Muse Spark
& 7 & 0.804 & Muse Spark & Meta \\
Kimi K2.6
& 8 & 0.801 & Kimi & Moonshot AI \\
GPT-5.2
& 9 & 0.795 & GPT & OpenAI \\
Qwen3.6 Plus
& 10 & 0.788 & Qwen & Alibaba Cloud / Qwen Team \\
Kimi K2.5
& 11 & 0.785 & Kimi & Moonshot AI \\
GPT-5
& 12 & 0.784 & GPT & OpenAI \\
MiniMax M3
& 13 & 0.781 & MiniMax & MiniMax \\
MiMo-V2.5
& 14 & 0.779 & MiMo & Xiaomi \\
Claude Opus 4.6
& 15 & 0.773 & Claude & Anthropic \\
Gemma 4 31B
& 16 & 0.769 & Gemma & Google \\
Qwen3.5-122B-A10B
& 16 & 0.769 & Qwen & Alibaba Cloud / Qwen Team \\
Gemini 3.1 Flash-Lite
& 18 & 0.768 & Gemini & Google \\
GPT-5.4 mini
& 19 & 0.766 & GPT & OpenAI \\
o3
& 20 & 0.764 & o-series & OpenAI \\
GPT-5.5 Instant
& 21 & 0.760 & GPT & OpenAI \\
Qwen3.6-27B
& 22 & 0.758 & Qwen & Alibaba Cloud / Qwen Team \\
Claude Sonnet 4.6
& 23 & 0.756 & Claude & Anthropic \\
Qwen3.6-35B-A3B
& 24 & 0.753 & Qwen & Alibaba Cloud / Qwen Team \\
Qwen3.5-35B-A3B
& 25 & 0.751 & Qwen & Alibaba Cloud / Qwen Team \\
\bottomrule
\end{tabular}
}
\caption{Twenty-five-entry MMMU-Pro leaderboard excerpt used during model
selection. The table defines part of the broader candidate landscape; it does
not determine automatic inclusion in the main study.}
\label{tab:mmmu-pro-context}
\end{table}

\section{Arena Vision Leaderboard Context}

Table~\ref{tab:arena-vision-context} presents the Arena Vision Overall context.
It reinforces the relevance of Claude, Gemini, GPT, Qwen, and Muse Spark as
candidate families for a vision-language study.

\begin{table}[H]
\centering
\small
\resizebox{\textwidth}{!}{%
\begin{tabular}{l c c l l}
\toprule
\textbf{Model name} &
\textbf{Rank} &
\textbf{Score} &
\textbf{Family} &
\textbf{Company} \\
\midrule
claude-opus-4-7-thinking
& 1 & $1309 \pm 8$ & Claude Opus & Anthropic \\
claude-fable-5
& 2 & $1307 \pm 15$ & Claude & Anthropic \\
claude-opus-4-6-thinking
& 3 & $1301 \pm 8$ & Claude Opus & Anthropic \\
claude-opus-4-7
& 4 & $1301 \pm 8$ & Claude Opus & Anthropic \\
claude-opus-4-6
& 5 & $1297 \pm 7$ & Claude Opus & Anthropic \\
muse-spark
& 6 & $1295 \pm 10$ & Muse Spark & Meta \\
gemini-3-pro
& 7 & $1290 \pm 8$ & Gemini & Google \\
claude-opus-4-8-thinking
& 8 & $1289 \pm 11$ & Claude Opus & Anthropic \\
gpt-5.5-high
& 9 & $1284 \pm 8$ & GPT & OpenAI \\
gpt-5.4-high
& 10 & $1283 \pm 8$ & GPT & OpenAI \\
gpt-5.5
& 11 & $1283 \pm 8$ & GPT & OpenAI \\
claude-opus-4-8
& 12 & $1280 \pm 11$ & Claude Opus & Anthropic \\
gemini-3.1-pro-preview
& 13 & $1280 \pm 6$ & Gemini & Google \\
gpt-5.2-chat-latest-20260210
& 14 & $1279 \pm 7$ & GPT & OpenAI \\
claude-sonnet-4-6
& 15 & $1278 \pm 7$ & Claude Sonnet & Anthropic \\
gpt-5.5-instant
& 16 & $1276 \pm 9$ & GPT & OpenAI \\
gpt-5.4
& 17 & $1275 \pm 8$ & GPT & OpenAI \\
gemini-3-flash
& 18 & $1272 \pm 6$ & Gemini & Google \\
kimi-k2.6
& 19 & $1264 \pm 8$ & Kimi & Moonshot \\
qwen3.7-plus-preview
& 20 & $1263 \pm 11$ & Qwen & Alibaba \\
dola-seed-2.0-pro
& 21 & $1259 \pm 8$ & Seed & ByteDance \\
gemini-3-flash (thinking-minimal)
& 22 & $1256 \pm 6$ & Gemini & Google \\
gpt-5.4-mini-high
& 23 & $1256 \pm 7$ & GPT & OpenAI \\
gemma-4-31b
& 24 & $1252 \pm 7$ & Gemma & Google \\
grok-4.20-beta-0309-reasoning
& 25 & $1251 \pm 7$ & Grok & xAI \\
\bottomrule
\end{tabular}
}
\caption{Twenty-five-entry Arena Vision Overall leaderboard excerpt used during
model selection. The table documents the broader candidate landscape; final
inclusion still requires completed repository artifacts and compatibility with
the common evaluation procedure.}
\label{tab:arena-vision-context}
\end{table}

\section{From Leaderboard Families to Final Cohort}

The final cohort is not a top-(k) slice of either leaderboard. It is the
intersection of three constraints: relevance in the broader model landscape,
compatibility with the common benchmark pipeline, and the existence of
completed repository artifacts.

The family-level screening proceeds as follows. The leaderboards identify five
main candidate families for this study: GPT, Gemini, Claude, Qwen, and Muse
Spark. Muse Spark is excluded despite its high rank because it is not generally
available through a reproducible public API route. Meta describes Muse Spark as
available at \url{meta.ai} and the Meta AI app, with API access offered only as
a private preview to selected users.\footnote{\url{https://ai.meta.com/blog/introducing-muse-spark-msl/}, accessed 2026-06-11.}
Artificial Analysis likewise reports that there are no currently benchmarked
API providers for Muse Spark.\footnote{\url{https://artificialanalysis.ai/models/muse-spark/providers}, accessed 2026-06-11.}

After excluding Muse Spark, the remaining families are GPT, Gemini, Claude, and
Qwen. GPT receives three slots because it is the strongest family for a
within-family analysis in this study: it appears repeatedly near the top of both
leaderboards and the repository contains completed artifacts for several GPT
generations. The other three families receive one model each. This keeps the
comparison broad across providers while preserving a controlled GPT-family
trajectory.

This gives the following six-model cohort:

\begin{table}[H]
\centering
\small
\begin{tabularx}{\textwidth}{@{}l l X@{}}
\toprule
\textbf{Model} & \textbf{Provider} & \textbf{Reason for inclusion} \\
\midrule
GPT-4o
& OpenAI
& Serves as the established GPT-family multimodal reference point and anchors
the within-family comparison. \\

GPT-4.1
& OpenAI
& Provides an intermediate GPT generation, allowing changes across OpenAI
models to be studied across tasks rather than inferred from a single pair. \\

GPT-5
& OpenAI
& Provides the latest GPT model evaluated in the cohort and completes the
three-model GPT-family comparison. \\

Claude Sonnet 4.5
& Anthropic
& Represents the Claude family with a strong general-purpose multimodal model
for cross-provider comparison. \\

Gemini 3.5 Flash
& Google
& Represents the Gemini family with a high-performing and practically usable
model that is especially relevant given Gemini's strong leaderboard presence. \\

Qwen 3.5
& Alibaba Cloud
& Represents the Qwen family and adds a non-US provider family to the main
comparison. \\
\bottomrule
\end{tabularx}
\caption{Models included in the main comparison.}
\label{tab:included-models}
\end{table}

The cohort is intentionally asymmetric. Equal provider counts would remove the
GPT generational comparison, which is one of the useful internal controls of
the study. Instead, the cohort uses a (3+1+1+1) design: three GPT models for
depth, and one Claude, one Gemini, and one Qwen model for breadth.

\section{Why Additional Leaderboard Models Were Not Included}

Many models visible in the leaderboard context were not included in the main
cohort. This does not mean they are unimportant. It means that including them
would have changed the study from a controlled repository-backed comparison
into a broader leaderboard survey.

In particular, additional GPT, Claude, Gemini, Qwen, Kimi, Gemma, Grok, MiMo,
Seed, MiniMax, and related systems could have been considered. They were not
automatically included because the study requires completed artifacts from the
same benchmark pipeline and a comparison setting close enough for the reported
results to be interpreted together.

The final cohort should therefore be read as a deliberately measured subset of
a larger model landscape. The leaderboards establish family relevance; the
repository artifacts determine model inclusion.

\section{Fine-Tuning Case Study}

Qwen2.5-VL-3B is included separately in a Fashion-MNIST fine-tuning case study.
It is not treated as another member of the main model ranking. The purpose of
this experiment is different: it measures whether parameter-efficient
fine-tuning can improve a compact model on a defined visual classification
task.

The case study compares the same Qwen2.5-VL-3B base model before and after LoRA
fine-tuning. This paired design controls for model architecture and scale, so
the observed difference can be attributed more directly to task-specific
adaptation.

Keeping this case study separate is necessary for a fair interpretation. Its
model size, deployment method, training exposure, and task specialization
differ from those of the general-purpose models in the main cohort. Its results
therefore support conclusions about fine-tuning effectiveness on Fashion-MNIST,
not a general claim that Qwen2.5-VL-3B is directly superior or inferior to the
six-model comparison cohort.

\section{Scope of the Conclusions}

The selected models support three forms of analysis:

\begin{itemize}
\item comparison of general-purpose multimodal systems across OpenAI,
Anthropic, Google, and Alibaba Cloud;
\item comparison of successive GPT-family models under the same benchmark
protocol;
\item comparison of a compact base model with its task-specific fine-tuned
counterpart in a separate case study.
\end{itemize}

Accordingly, conclusions are limited to the evaluated model versions,
endpoints, prompts, benchmark samples, and execution settings. The leaderboard
tables document the broader model landscape and justify why the selected
families were reasonable candidates. They do not override the stricter
inclusion rule that reported comparisons must be traceable to completed
artifacts produced by the common evaluation pipeline.

\end{document}
